---
title: "Automatic Construction of Bootstrap Confidence Intervals"
author: "Bradley Efron and Balasubramanian Narasimhan"
date: '`r Sys.Date()`'
bibliography: bcaboot.bib
output:
  html_document:
  fig_caption: yes
  theme: cerulean
  toc: yes
  toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Automatic Construction of Bootstrap Confidence Intervals}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r echo=FALSE}
knitr::opts_chunk$set(
    message = FALSE,
    warning = FALSE,
    error = FALSE,
    tidy = FALSE,
    cache = FALSE
)
library(bcaboot)
```

## Introduction

Bootstrap confidence intervals depend on three elements:

- the cdf of the $B$ bootstrap replications $t_i^*$, $i=1\ldots B$
- the bias-correction number $z_0 = \Phi(\sum_i^B I(t_i^* < t_0) / B )$
  where $t_0=f(x)$ is the original estimate
- the acceleration number $a$ that measures the rate of
  change in $\sigma_{t_0}$ as $x$, the data changes.

The first two of these depend only on the bootstrap distribution, and
not how it is generated: parametrically or non-parametrically.

Package `bcaboot` aims to make construction of bootstrap confidence
intervals _almost_ automatic. The two main functions are:

- `bca_nonpar()` for nonparametric bootstrap
- `bca_par()` for parametric bootstrap

All results are returned as `bcaboot` objects with a consistent
structure and support `tidy()`, `glance()`, and `autoplot()` for
integration with the tidyverse ecosystem.

Further details are in the @efronnaras2018 paper. Much of the theory
behind the approach can be found in references @efron1987,
@diciccio1992, @diciccio1996, and @efron2016.

## A Nonparametric Example

Suppose we wish to construct bootstrap confidence intervals for an
$R^2$-statistic from a linear regression. Using the diabetes data from
the [`lars`](https://cran.r-project.org/package=lars) (442 by 11) as
an example, we use the function below to regress the `y` on `x`, a
matrix of of 10 predictors, to compute $R^2$.

```{r}
data(diabetes, package = "bcaboot")
Xy <- cbind(diabetes$x, diabetes$y)
rfun <- function(Xy) {
    y <- Xy[, 11]
    X <- Xy[, 1:10]
    summary(lm(y ~ X) )$adj.r.squared
}
```

Constructing bootstrap confidence intervals involves merely calling
`bca_nonpar`:

```{r}
set.seed(1234)
result <- bca_nonpar(x = Xy, B = 2000, func = rfun, verbose = FALSE)
```

The `result` prints a clean summary:

```{r}
result
```

### Tidy output

The `tidy()` method returns a tibble with one row per (confidence
level, method) combination, following the broom conventions:

```{r}
tidy(result)
```

The `glance()` method provides a one-row summary of the bootstrap run:

```{r}
glance(result)
```

### Visualization

When `ggplot2` is available, `autoplot()` produces a publication-ready
plot of the confidence intervals:

```{r fig.width = 7, fig.height = 5}
library(ggplot2)
autoplot(result)
```

### Acceleration methods

The `accel` argument controls how the acceleration $a$ is estimated:

- `accel = "regression"` (default): uses local regression on bootstrap
  count vectors nearest to uniform. Also computes the gbca diagnostic.
- `accel = "jackknife"`: uses classical delete-one (or delete-group)
  jackknife influence values.

```{r}
set.seed(1234)
result_jk <- bca_nonpar(x = Xy, B = 2000, func = rfun,
                        accel = "jackknife", verbose = FALSE)
tidy(result_jk)
```

## A Parametric Example

```{r, echo = FALSE}
if (!requireNamespace("glmnet", quietly = TRUE)) {
    stop("Please install glmnet package for this vignette.")
}
load(system.file("extdata", "neonates.rda", package = "bcaboot"))
```

A logistic regression was fit to data on 812 neonates at a large
clinic. Here is a summary of the dataset.

```{r}
str(neonates)
```

The goal was to predict death versus survival---$y$ is 1 or 0,
respectively---on the basis of 11 baseline variables of which one of
them `resp` was of particular concern. (There were 207 deaths and 605
survivors.) So here $\theta$, the parameter of interest is the
coefficient of `resp`. Discussions with the investigator suggested a
weighting of 4 to 1 of deaths versus non-deaths.

### A Logistic Model

```{r}
weights <- with(neonates, ifelse(y == 0, 1, 4))
glm_model <- glm(formula = y ~ ., family = "binomial", weights = weights, data = neonates)
summary(glm_model)
```
Parametric bootstrapping in this context requires us to independently
sample the response according to the estimated probabilities from
regression model. As discussed in the paper accompanying this
software, routine `bca_par` also requires sufficient statistics
$\hat{\beta} = M^\prime y$ where $M$ is the model matrix. Therefore,
it makes sense to have a function do the work. The function `glm_boot`
below returns a list of the estimate $\hat{\theta}$, the bootstrap
estimates, and the sufficient statistics.

```{r}
glm_boot <- function(B, glm_model, weights, var = "resp") {
    pi_hat <- glm_model$fitted.values
    n <- length(pi_hat)
    y_star <- sapply(seq_len(B), function(i) ifelse(runif(n) <= pi_hat, 1, 0))
    beta_star <- apply(y_star, 2, function(y) {
        boot_data <- glm_model$data
        boot_data$y <- y
        coef(glm(formula = y ~ ., data = boot_data, weights = weights, family = "binomial"))
    })
    list(theta = coef(glm_model)[var],
         theta_star = beta_star[var, ],
         suff_stat = t(y_star) %*% model.matrix(glm_model))
}
```

Now we can compute the bootstrap estimates using `bca_par`.

```{r}
set.seed(3891)
glm_boot_out <- glm_boot(B = 2000, glm_model = glm_model, weights = weights)
glm_bca <- bca_par(t0 = glm_boot_out$theta,
                   tt = glm_boot_out$theta_star,
                   bb = glm_boot_out$suff_stat)
```

We can examine the results using tidy methods:

```{r}
glm_bca
```

```{r}
tidy(glm_bca)
```

Our bootstrap standard error using $B=2000$ samples for `resp` can be
read off from the glance output:

```{r}
glance(glm_bca)
```


### A Penalized Logistic Model

Now suppose we wish to use a nonstandard estimation procedure, for
example, via the `glmnet` package, which uses cross-validation to
figure out a best fit, corresponding to a penalization parameter
$\lambda$ (named `lambda.min`).

```{r}
X <- as.matrix(neonates[, seq_len(11)]) ; Y <- neonates$y;
glmnet_model <- glmnet::cv.glmnet(x = X, y = Y, family = "binomial", weights = weights)
```

We can examine the estimates at the `lambda.min` as follows.

```{r}
coefs <- as.matrix(coef(glmnet_model, s = glmnet_model$lambda.min))
knitr::kable(data.frame(variable = rownames(coefs), coefficient = coefs[, 1]), row.names = FALSE, digits = 3)
```

Following the lines above, we create a helper function to perform the
bootstrap.

```{r}
glmnet_boot <- function(B, X, y, glmnet_model, weights, var = "resp") {
    lambda <- glmnet_model$lambda.min
    theta <- as.matrix(coef(glmnet_model, s = lambda))
    pi_hat <- predict(glmnet_model, newx = X, s = "lambda.min", type = "response")
    n <- length(pi_hat)
    y_star <- sapply(seq_len(B), function(i) ifelse(runif(n) <= pi_hat, 1, 0))
    beta_star <- apply(y_star, 2,
                       function(y) {
                           as.matrix(coef(glmnet::glmnet(x = X, y = y, lambda = lambda, weights = weights, family = "binomial")))
                       })

    rownames(beta_star) <- rownames(theta)
    list(theta = theta[var, ],
         theta_star = beta_star[var, ],
         suff_stat = t(y_star) %*% X)
}
```

And off we go.

```{r}
glmnet_boot_out <- glmnet_boot(B = 2000, X, y, glmnet_model, weights)
glmnet_bca <- bca_par(t0 = glmnet_boot_out$theta,
                      tt = glmnet_boot_out$theta_star,
                      bb = glmnet_boot_out$suff_stat)
```

We can compare both models side-by-side:

```{r fig.width = 7, fig.height = 5}
autoplot(glm_bca) + ggtitle("glm model")
```

```{r fig.width = 7, fig.height = 5}
autoplot(glmnet_bca) + ggtitle("glmnet model")
```

## Ratio of Independent Variance Estimates

Assume we have two independent estimates of variance from normal
theory:

$$
\hat{\sigma}_1^2\sim\frac{\sigma_1^2\chi_{n_1}^2}{n_1},
$$

and

$$
\hat{\sigma}_2^2\sim\frac{\sigma_2^2\chi_{n_2}^2}{n_2}.
$$

Suppose now that our parameter of interest is

$$
	\theta=\frac{\sigma_1^2}{\sigma_2^2}
$$

for which we wish to compute confidence limits. In this setting, theory yields exact limits:

$$
\hat{\theta}(\alpha) = \frac{\hat{\theta}}{F_{n_1,n_2}^{1-\alpha}}.
$$

We can apply `bca_par` to this problem. As before, here are our helper
functions.

```{r}
ratio_boot <- function(B, v1, v2) {
    s1 <- sqrt(v1) * rchisq(n = B, df = n1)  / n1
    s2 <- sqrt(v2) * rchisq(n = B, df = n2)  / n2
    theta_star <- s1 / s2
    beta_star <- cbind(s1, s2)
    list(theta = v1 / v2,
         theta_star = theta_star,
         suff_stat = beta_star)
}

funcF <- function(beta) {
    beta[1] / beta[2]
}
```

Note that we have an additional function `funcF` which corresponds to
$\tau(\hat{\beta}^*)$ in the paper. This is the function expressing
the parameter of interest as a function of the sample.

```{r}
B <- 16000; n1 <- 10; n2 <- 42
ratio_boot_out <- ratio_boot(B, 1, 1)
ratio_bca <- bca_par(t0 = ratio_boot_out$theta,
                     tt = ratio_boot_out$theta_star,
                     bb = ratio_boot_out$suff_stat, func = funcF)
```

The tidy output shows the BCa, standard, and ABC limits:

```{r}
tidy(ratio_bca)
```

We can compare against exact F-distribution limits:

```{r}
exact <- 1 / qf(df1 = n1, df2 = n2, p = 1 - as.numeric(rownames(ratio_bca$limits)))
knitr::kable(cbind(ratio_bca$limits, exact = exact), digits = 3)
```

Clearly the BCa limits match the exact values very well and suggests a
large upward correction to the standard limits.

```{r fig.width = 7, fig.height = 5}
autoplot(ratio_bca)
```

## Migration from pre-1.0 API {#migration}

Version 1.0 introduces a new API. The old functions continue to work
but emit deprecation warnings. Here is how to migrate:

### Function mapping

| Old function | New function | Notes |
|:-------------|:-------------|:------|
| `bcajack(x, B, func)` | `bca_nonpar(x, B, func, accel = "jackknife")` | Classical jackknife acceleration |
| `bcajack2(x, B, func)` | `bca_nonpar(x, B, func, accel = "regression")` | Regression acceleration (default) |
| `bcanon(B, x, func)` | `bca_nonpar(x, B, func)` | Removed; note argument order change |
| `bcapar(t0, tt, bb)` | `bca_par(t0, tt, bb)` | |
| `bcaplot(result)` | `autoplot(result)` | Requires ggplot2 |

### Parameter mapping

| Old | New | Description |
|:----|:----|:------------|
| `alpha` | `conf.level` | E.g., `alpha = 0.025` becomes `conf.level = 0.95` |
| `m` | `n_groups` | Number of jackknife groups |
| `mr` | `group_reps` | Random grouping repetitions |
| `pct` | `kl_fraction` | Fraction of nearby count vectors |
| `K` | `n_jack` | Delete-d jackknife repetitions |
| `J` | `jack_groups` | Groups per jackknife fold |
| `trun` | `truncation` | Truncation for acceleration |
| `cd` | `conf_density` | Confidence density flag |
| `B` (as list) | `boot_data` | Pre-computed bootstrap data |

### Return structure

The new objects use `$limits` (not `$lims`), `$B_mean` (not
`$B.mean`), and `$stats` is always a 2-row matrix (never a named
vector). Use `tidy()` and `glance()` instead of accessing list
components directly:

```r
## Old way
result$lims[, "bca"]
result$stats["theta"]

## New way
tidy(result)    # tibble of confidence limits
glance(result)  # tibble of summary statistics
```

### Deprecation timeline

| Release | Status |
|:--------|:-------|
| 1.0 | Old functions work with once-per-session warning |
| 1.1 | Old functions error with migration message |
| 2.0 | Old functions removed |

## References

